\newpage
\phantomsection
\addcontentsline{toc}{section}{REFERENCES}

\bibliographystyle{IEEEtran}

\begin{thebibliography}{25}

\bibitem{react_docs}
Meta Platforms, Inc., ``React --- A JavaScript Library for Building User Interfaces,'' \textit{React Documentation}, 2024. [Online]. Available: https://react.dev/

\bibitem{django_docs}
Django Software Foundation, ``Django: The Web Framework for Perfectionists with Deadlines,'' \textit{Django Documentation}, 2024. [Online]. Available: https://docs.djangoproject.com/

\bibitem{drf_docs}
T. Christie et al., ``Django REST Framework,'' \textit{DRF Documentation}, 2024. [Online]. Available: https://www.django-rest-framework.org/

\bibitem{mui_docs}
MUI (formerly Material-UI), ``MUI: The React Component Library,'' \textit{MUI Documentation}, 2024. [Online]. Available: https://mui.com/

\bibitem{jwt_rfc}
M.~Jones, J.~Bradley, and N.~Sakimura, ``JSON Web Token (JWT),'' \textit{RFC 7519}, Internet Engineering Task Force, May 2015. [Online]. Available: https://tools.ietf.org/html/rfc7519

\bibitem{simplejwt}
D.~Alling, ``Simple JWT --- A JSON Web Token Authentication Plugin for Django REST Framework,'' 2024. [Online]. Available: https://django-rest-framework-simplejwt.readthedocs.io/

\bibitem{capacitor_docs}
Ionic, ``Capacitor: Cross-Platform Native Runtime for Web Apps,'' \textit{Capacitor Documentation}, 2024. [Online]. Available: https://capacitorjs.com/docs

\bibitem{vite_docs}
E.~You et al., ``Vite: Next Generation Frontend Tooling,'' \textit{Vite Documentation}, 2024. [Online]. Available: https://vitejs.dev/

\bibitem{axios_docs}
M.~Zabriskie et al., ``Axios: Promise-Based HTTP Client for the Browser and Node.js,'' 2024. [Online]. Available: https://axios-http.com/

\bibitem{recharts_docs}
Recharts Team, ``Recharts: A Composable Charting Library Built on React Components,'' 2024. [Online]. Available: https://recharts.org/

\bibitem{fullcalendar_docs}
A.~Shaw, ``FullCalendar: Full-Sized Drag \& Drop JavaScript Event Calendar,'' 2024. [Online]. Available: https://fullcalendar.io/

\bibitem{framer_motion}
Framer, ``Framer Motion: A Production-Ready Motion Library for React,'' 2024. [Online]. Available: https://www.framer.com/motion/

\bibitem{gemini_api}
Google DeepMind, ``Gemini API Documentation,'' \textit{Google AI for Developers}, 2024. [Online]. Available: https://ai.google.dev/docs

\bibitem{react_router}
Remix Software, ``React Router: Declarative Routing for React,'' 2024. [Online]. Available: https://reactrouter.com/

\bibitem{react_query}
T.~Linsley, ``TanStack Query: Powerful Asynchronous State Management for TS/JS, React, Solid, Vue and Svelte,'' 2024. [Online]. Available: https://tanstack.com/query

\bibitem{html5qrcode}
M.~Sharma, ``html5-qrcode: A Cross Platform HTML5 QR Code \& Barcode Reader,'' 2024. [Online]. Available: https://github.com/mebjas/html5-qrcode

\bibitem{reportlab}
ReportLab Inc., ``ReportLab: The Proven Open Source Engine for Creating Complex, Data-Rich PDF Documents,'' 2024. [Online]. Available: https://www.reportlab.com/

\bibitem{cors_headers}
A.~Johnson, ``django-cors-headers: A Django App for Handling the Server Headers Required for Cross-Origin Resource Sharing,'' 2024. [Online]. Available: https://github.com/adamchainz/django-cors-headers

\bibitem{pillow_docs}
A.~Clark et al., ``Pillow: The Friendly PIL Fork,'' 2024. [Online]. Available: https://pillow.readthedocs.io/

\bibitem{gunicorn_docs}
B.~Chesneau et al., ``Gunicorn: Python WSGI HTTP Server for UNIX,'' 2024. [Online]. Available: https://gunicorn.org/

\bibitem{agile_manifesto}
K.~Beck et al., ``Manifesto for Agile Software Development,'' 2001. [Online]. Available: https://agilemanifesto.org/

\bibitem{restful_api}
R.~T.~Fielding, ``Architectural Styles and the Design of Network-Based Software Architectures,'' Ph.D. dissertation, University of California, Irvine, 2000.

\bibitem{material_design}
Google LLC, ``Material Design Guidelines,'' 2024. [Online]. Available: https://material.io/design

\bibitem{oauth2}
D.~Hardt, ``The OAuth 2.0 Authorization Framework,'' \textit{RFC 6749}, Internet Engineering Task Force, Oct. 2012.

\bibitem{sqlite}
D.~R.~Hipp, ``SQLite: A Self-Contained, Serverless, Zero-Configuration, Transactional SQL Database Engine,'' 2024. [Online]. Available: https://www.sqlite.org/

\end{thebibliography}
